---
title: "DDD Final"
author: "Amouki Kaayou"
date: "1/16/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DDD Final Project(Amouki Kaayou)

## 1. Introduction

  Here is my Markdown Final Project for DDD (Winter 2021). The data analyzed is the "Yearbook of Immigration of 2019", a collection of excel files reporting numbers regarding immigration both for the Fiscal year 2019, but also throughout the years up until 2019. The entire yearbook consists of 41 tables which can be split in 5 categories
  * Lawful PErmanent Residents
  * Refugees and Asylees
  * Naturalization
  * Non-immigrants (Visa recipients)
  * 'Aliens' and Law Enforcement
  The data can be retrieved at the following link (https://www.dhs.gov/immigration-statistics/yearbook/2019)
I selected this data because I had wanted to use it in the past but was unable/too lazy to go through all files and edit them. 
  I will fist analyze the data spanning 19 years (from 2000 to 2019) in order to see if a change in administration (Republican vs Democrats) affects immigration numbers in any way. Then I will analyse some of the data in regard to numbers by country of provenance over the span of 10 years (2010 to 2019) to see if there has been any recent changes in immigration fluxes in regard to the last 2 presidents.


## 1. Required packages and loading data

```{r Package loading, eval=FALSE}

#There are the packages that will be needed to call the different functions used throughout the Mardkown

library(tidyverse)
library(readxl)
library(reader)
library(readr)
library(stringr)
library(lubridate)
library(ggplot2)
library(dplyr)
library(pander)
library(gt)

```

## 1. Data Cleaning 
###  1. Variable Selection and Creation

  After loading all tables, I viewed each one of them one by one and separated the tables that only pertained to the fiscal year as they would not offer a lot to analyse since all tables have strictly numeric values. After settling down on the datasets I wanted to use, I nested them in variables that would be easier to manipulate


```{r Variable Selection and Creation, eval=FALSE}

#Lawful Permanent Residents Data

lpr_genX <- read_excel("fy2019_table1.xlsx")
lpr_ByLastRegionResidenceX <- read_excel("fy2019_table2.xlsx")
lpr_ByRegionBirthX <- read_excel("fy2019_table3d.xlsx")
lpr_ByRegionResidence <- read_excel("fy2019_table4.xlsx")
lpr_ByCBSAREsidence <- read_excel("fy2019_table5.xlsx")
lpr_ByMajorClass <- read_excel("fy2019_table6.xlsx")


#Refugees and Asylees Data

ref_arrivals <- read_excel("fy2019_table13.xlsx")
ref_arr_ByNationality <- read_excel("fy2019_table14d.xlsx")
ref_asyl_grantedAD <- read_excel("fy2019_table16.xlsx")
ref_asyl_grantedA_ByNationality <- read_excel("fy2019_table17d.xlsx")
ref_asyl_grantedD_ByNationality <- read_excel("fy2019_table19d.xlsx")


#Naturalization Data

nat_gen <- read_excel("fy2019_table20.xlsx")
nat_ByRegionBirth <- read_excel("fy2019_table21d.xlsx")
nat_ByRegionResidence <- read_excel("fy2019_table22.xlsx")
nat_ByCBSAResidence <- read_excel("fy2019_table23.xlsx")


#Non-immigrants (Visa Recipients) Data

nonimg_ByClass <- read_excel("fy2019_table25d.xlsx")
nonimg_ByRegionCitizenship <- read_excel("fy2019_table26.xlsx")
nonimg_ByRegionResidence <- read_excel("fy2019_table27d.xlsx")

#'Aliens' and Law Enforcement Data

aliens_appreh <- aliens_gen <- read_excel("fy2019_table33.xlsx")
aliens_appreh_ByRegionNat  <- read_excel("fy2019_table34d.xlsx")
aliens_appreh_ByJuris <- read_excel("fy2019_table35.xlsx")
aliens_inadmin <- read_excel("fy2019_table36.xlsx")
aliens_inadmin_ByRegionNat <- read_excel("fy2019_table37d.xlsx")
aliens_inadmin_ByFieldOffice <- read_excel("fy2019_table38.xlsx")
aliens_remov <- read_excel("fy2019_table39.xlsx")
aliens_remov_ByRegionNat <- read_excel("fy2019_table40d.xlsx")
aliens_remov_byCrimS_RegionNat <- read_excel("fy2019_table41d.xlsx")

```

###  1. Data Cleaning: By year
  
  After that I started working on the data, I first too kthe data sets that had data spanning over 20 years and started cleaning them. This si the data I wil use to build graphs over time. 

```{r Data Cleaning for datasets with years, eval=FALSE}

#Lawful Permanent Residents Data
#LPR Status granted from 2000 to 2019
lpr_genX <- read_excel("fy2019_table1.xlsx")
lpr1 <- lpr_genX
lpr1 <- lpr1[-c(1:33,54,55), -c(1:6)]
colnames(lpr1) <- c("Year","People Obtaining LPR Status")
lpr1$`People Obtaining LPR Status` <- as.numeric(lpr1$`People Obtaining LPR Status`)
LPR_years <- lpr1


#Refugees and Asylees Data
#Refugee Arrivals from 2000 to 2019
ref_arrivals <- read_excel("fy2019_table13.xlsx")
ref1 <- ref_arrivals
colnames(ref1) <- c("Year","Refugee Arrivals")
ref1$`Refugee Arrivals` <- as.numeric(ref1$`Refugee Arrivals`)
ref1 <- ref1[-c(1:23,44,45), ]

#Asylums granted from 2000 to 2019
ref_asyl_grantedAD <- read_excel("fy2019_table16.xlsx")
ref3 <- ref_asyl_grantedAD
colnames(ref3) <- c("Year","Total Asylums Granted","Asylums Granted Affirmatvely","Asylums Granted Defensively")
ref3$`Total Asylums Granted` <- as.numeric(ref3$`Total Asylums Granted`)
ref3$`Asylums Granted Affirmatvely` <- as.numeric(ref3$`Asylums Granted Affirmatvely`)
ref3$`Asylums Granted Defensively` <- as.numeric(ref3$`Asylums Granted Defensively`)
ref3 <- ref3[-c(1:13,34),]


#Naturalization Data
##Naturalization petitions filled, granted and denied from 2000 to 2019
nat_gen <- read_excel("fy2019_table20.xlsx")
nat1 <- nat_gen
colnames(nat1) <- c("Year","Petitions Filled"," Total Persons Naturalized","Civilians Naturalized","Military Naturalized", "Not Reported","Petitions Denied")
nat1 <- nat1[-c(1:98,119:124),]
NAT_years <- nat1


#'Aliens' and Law Enforcement Data
#'Aliens' apprehended from 2000 to 2019
aliens_appreh <- aliens_gen <- read_excel("fy2019_table33.xlsx")
enf1 <- aliens_appreh
colnames(enf1) <- c("X","Y","Year","Aliens Apprehended")
enf1 <- enf1[-c(1:22,43:65), -c(1,2)]
enf1$Year[9]=2008
enf1$Year[10]=2009
enf1$Year[17]=2016

#'Aliens' deemed inadmissible from 2000 to 2019
aliens_inadmin <- read_excel("fy2019_table36.xlsx")
enf4 <- aliens_inadmin
colnames(enf4) <- c("Year","Aliens Determined Inadmissible")
enf4 <- enf4[-c(1:3,19,20),]
enf4.new <- data.frame(c(2000, 2001, 2002, 2003, 2004),c("NA", "NA", "NA", "NA","NA"))
colnames(enf4.new) <- c("Year","Aliens Determined Inadmissible")
enf4 <- rbind(enf4,enf4.new)
enf4 <- enf4 %>% dplyr::arrange(Year)

#'Aliens' removed from 2000 to 2019
aliens_remov <- read_excel("fy2019_table39.xlsx")
enf7 <-aliens_remov
colnames(enf7) <- c("Year","Aliens Removals","Aliens Returned")
enf7 <- enf7[-c(1:111,132:137),]
enf7$Year[17]=2016

```

  After the datasets were cleaned, I looked at those that were in the same area and merged them so the data would be centralized.

```{r Data merging for datsets with years, eval=FALSE}
#Refugees and Asylees Data
#Merging data on Refugees and Asylums for the 2000-2019 period
REF_years <- merge(ref1, ref3, by = "Year")


#'Aliens' and Law Enforcement Data
#Merging data on 'aliens' for the 2000-2019 period
ENF_years <- merge(enf1, enf4,by = "Year")
ENF_years <- merge(ENF_years, enf7, by = "Year")
ENF_years$`Aliens Apprehended` <- as.numeric(ENF_years$`Aliens Apprehended`)
ENF_years$`Aliens Determined Inadmissible` <- as.numeric(ENF_years$`Aliens Determined Inadmissible`)
ENF_years$`Aliens Removals` <- as.numeric(ENF_years$`Aliens Removals`)
ENF_years$`Aliens Returned` <- as.numeric(ENF_years$`Aliens Returned`)

```

###  1. Data Cleaning: By State
  
  After I started working on the datasets that included data in relation of years (2010 to 2019) and regions, I sorted those datasets in two groups. We first have one small group of datasets that were about Region of Residence and only included US States:

```{r Data Cleaning for datasets with states, eval=FALSE}

#Lawful Permanent Residents Data
#LPR Status Granted by State of Residency from 2010 to 2019
#This code sorts the data by State
lpr_ByRegionResidence <- read_excel("fy2019_table4.xlsx")
lpr4 <- lpr_ByRegionResidence
lpr4 <- lpr4[-c(1,2,60:62), ]
lpr4t <- lpr4
lpr4t <- lpr4t[-c(1,2,60:62), ]
names_lpr4t <- lpr4t[1,]
colnames(lpr4t) <- names_lpr4t
lpr4t <- lpr4t[-c(1),]
LPR_state_table <- lpr4t
#This code sorts the data by Year
lpr_ByRegionResidence <- read_excel("fy2019_table4.xlsx")
lpr4x <- lpr_ByRegionResidence
lpr4x <- lpr4x[-c(1,2,60:62), ]
names1 <- lpr4x$`Table 4.`
lpr4x <- lpr4x[,-c(1)]
lpr4x <- t(lpr4x)
colnames(lpr4x) <- names1
colnames(lpr4x)[1] <- "Year"
colnames(lpr4x)[56] <- "Other"
lpr4x[lpr4x=="-"]<-0


#Naturalization Data
#Naturalization granted by State of Residence from 2010 to 2019
#This code sorts the data by State
nat_ByRegionResidence <- read_excel("fy2019_table22.xlsx")
nat3 <- nat_ByRegionResidence
nat3 <- nat3[-c(1,2,60:62), ]
nat3t <- nat3
names_nat3t <- nat3t[1,]
colnames(nat3t) <- names_nat3t
nat3t <- nat3t[-c(1),]
NAT_state_table <- nat3t
#This code sorts the data by Year
nat_ByRegionResidence <- read_excel("fy2019_table22.xlsx")
nat3x <- nat_ByRegionResidence
nat3x <- nat3[-c(1,2,60:62), ]
names2 <- nat3x$`Table 22.`
nat3x <- nat3x[,-c(1)]
nat3x <- t(nat3x)
colnames(nat3x) <- names2
colnames(nat3x)[1] <- "Year"
colnames(lpr4x)[56] <- "Other"

```

###  1. Data Cleaning: By Country/Continent
  The second group includes the datasets that included either Region of Nationality or Provenance, which included both Continents and/or countries.

```{r Data Cleaning for satasets with Countries and Continents, eval=FALSE}

#Lawful Permanent Residents Data
#LPR Status by Last Country of Residence from 2010 to 2019
#This code sorts it by Year
lpr_ByLastRegionResidenceX <- read_excel("fy2019_table2.xlsx")
lpr2x <- lpr_ByLastRegionResidenceX
lpr2x <- lpr2x[-c(1:2,92:115), -c(2:20)]
names <- lpr2x$`Table 2.`
lpr2x <- lpr2x[,-c(1)]
lpr2x <- t(lpr2x)
colnames(lpr2x) <- names
colnames(lpr2x)[1] <- "Year"
lpr2x <- lpr2x[,-c(4,18,62)]
colnames(lpr2x)[4] <- "Austria"
colnames(lpr2x)[5] <- "Hungary"
colnames(lpr2x)[17] <- "Norway"
colnames(lpr2x)[18] <- "Sweden"
lpr2x[lpr2x=="-"]<-0
#This code sorts it by Regions
lpr_ByLastRegionResidenceX <- read_excel("fy2019_table2.xlsx")
lpr2 <- lpr_ByLastRegionResidenceX
lpr2 <- lpr2[-c(1:2,92:115), -c(2:20)]
lpr2t <- lpr2
names_lpr2t <- lpr2t[1,]
colnames(lpr2t) <- names_lpr2t
lpr2t <- lpr2t[-c(1),]
lpr2t[4,1] <- "Austria"
lpr2t[5,1] <- "Hungary"
lpr2t[18,1] <- "Norway"
lpr2t[19,1] <- "Sweden"
lpr2t <- lpr2t[-c(3,17),]
lpr2t[lpr2t=="X"]<-"NA"
lpr2t[lpr2t=="D"]<-"witheld"
lpr2t[lpr2t=="-"]<-"0"
as.factor(lpr2t$`Region and country of last residence`)
levels(lpr2t$`Region and country of last residence`) <- c("Europe","Asia","America","Central America","South America","Africa","Oceania")
lpr2t$`Region and country of last residence`[3:26] <- "Europe"
lpr2t$`Region and country of last residence`[28:41] <- "Asia"
lpr2t$`Region and country of last residence`[43:50] <- "America"
lpr2t$`Region and country of last residence`[52:59] <- "Central America"
lpr2t$`Region and country of last residence`[61:73] <- "South America"
lpr2t$`Region and country of last residence`[74] <- "America"
lpr2t$`Region and country of last residence`[76:81] <- "Africa"
lpr2t$`Region and country of last residence`[83:85] <- "Oceania"
LPR_LastResidence_table <- lpr2t
#This code sorts it by continent
LPR_REGION_continent <- LPR_REGION_table %>% slice(1,2,27,42,51,60,75,82)
#This code sorts it by country
LPR_REGION_country <- LPR_REGION_table %>% slice(3:26,28:41,43:50,52:59,61:74,76:81,83:85)

#LPR Status by Country of Birth from 2010 to 2019
#This code sorts it by Year
lpr_ByRegionBirthX <- read_excel("fy2019_table3d.xlsx")
lpr3x <- lpr_ByRegionBirthX
lpr3x <- lpr3x[-c(1:2,219:224),]
names3 <- lpr3x$`Table 3.`
lpr3x <- lpr3x[,-c(1)]
lpr3x <- t(lpr3x)
colnames(lpr3x) <- names3
colnames(lpr3x)[1] <- "Year"
lpr3x <- lpr3x[,-c(2,11)]
lpr3x[lpr3x=="X"]<-"NA"
lpr3x[lpr3x=="D"]<-"witheld"
lpr3x[lpr3x=="-"]<-0
#This code sorts it by Regions
lpr_ByRegionBirthX <- read_excel("fy2019_table3d.xlsx")
lpr3t <- lpr_ByRegionBirthX
lpr3t <- lpr3t[-c(1:2,219:224),]
names_lpr3t <- lpr3t[1,]
colnames(lpr3t) <- names_lpr3t
lpr3t <- lpr3t[-c(1),]
lpr3t <- lpr3t[-c(1,10),]
lpr3t[lpr3t=="X"]<-"NA"
lpr3t[lpr3t=="D"]<-"witheld"
lpr3t[lpr3t=="-"]<-"0"
lpr3t[1,1] <- "Total Continent"
lpr3t[9,1] <- "Total Country"
as.factor(lpr3t$`Region and country of birth`)
levels(lpr3t$`Region and country of birth`) <- c("Continent","Country")
lpr3t$`Region and country of birth`[1:8] <- "Continent"
lpr3t$`Region and country of birth`[9:213] <- "Country"
LPR_Birth_table <- lpr3t
#This code sorts it by continent
LPR_Birth_continent <- LPR_Birth_table %>% slice(1:8)
#This code sorts it by country
LPR_Birth_country <- LPR_Birth_table %>% slice(9:213)


#Refugees and Asylees Data
#Refugee Arrivals by Nationality from 2010 to 2019
#This code sorts it by years
ref_arr_ByNationality <- read_excel("fy2019_table14d.xlsx")
ref2x <- ref_arr_ByNationality
ref2x <- ref2x[-c(1,2,4,10,76:82), ]
names_ref2x <- ref2x$`Table 14.`
ref2x <- ref2x[,-c(1)]
ref2x <- t(ref2x)
colnames(ref2x) <- names_ref2x
colnames(ref2x)[1] <- "Year"
colnames(ref2x)[70] <- "All Other Countries"
colnames(ref2x)[71] <- "Unknown"
ref2x[ref2x=="X"]<-"NA"
ref2x[ref2x=="D"]<-"witheld"
ref2x[ref2x=="-"]<-0
#This code sorts it by Regions
ref_arr_ByNationality <- read_excel("fy2019_table14d.xlsx")
ref2t <- ref_arr_ByNationality
ref2t <- ref2t[-c(1,2,4,10,76:82), ]
names_ref2t <- ref2t[1,]
colnames(ref2t) <- names_ref2t
ref2t <- ref2t[-c(1,9),]
ref2t[ref2t=="X"]<-"NA"
ref2t[ref2t=="D"]<-"witheld"
ref2t[ref2t=="-"]<-"0"
ref2t[1,1] <- "Total Continent"
ref2t[9,1] <- "Total Country"
ref2t <- ref2t[-c(8,9), ]
as.factor(ref2t$`Region and country of nationality`)
levels(ref2t$`Region and country of nationality`) <- c("Continent","Country")
ref2t$`Region and country of nationality`[1:7] <- "Continent"
ref2t$`Region and country of nationality`[8:69] <- "Country"
REF_Nationality_table <- ref2t
#This code sorts it by continent
REF_Nat_continent <- REF_Nationality_table %>% slice(1:7)
#This code sorts it by country
REF_Nat_country <- REF_Nationality_table %>% slice(8:69)

#Affirmative Asylums granted by Nationality from 2010 to 2019
#This code sorts it by years
ref_asyl_grantedA_ByNationality <- read_excel("fy2019_table17d.xlsx")
ref4x <- ref_asyl_grantedA_ByNationality
ref4x <- ref4x[-c(1,2,4,13,121:125), ]
names_ref4x <- ref4x$`Table 17.`
ref4x <- ref4x[,-c(1)]
ref4x <- t(ref4x)
colnames(ref4x) <- names_ref4x
colnames(ref4x)[1] <- "Year"
colnames(ref4x)[115] <- "All Other Countries"
ref4x[ref4x=="X"]<-"NA"
ref4x[ref4x=="D"]<-"witheld"
ref4x[ref4x=="-"]<-0
#This code sorts it by Regions
ref_asyl_grantedA_ByNationality <- read_excel("fy2019_table17d.xlsx")
ref4t <- ref_asyl_grantedA_ByNationality
ref4t <- ref4t[-c(1,2,4,13,121:125), ]
names_ref4t <- ref4t[1,]
colnames(ref4t) <- names_ref4t
ref4t <- ref4t[-c(1),]
ref4t[ref4t=="X"]<-"NA"
ref4t[ref4t=="D"]<-"witheld"
ref4t[ref4t=="-"]<-"0"
ref4t[1,1] <- "Total Continent"
ref4t[9,1] <- "Total Country"
REF_AsyA_table <- ref4t
#This code sorts it by continent
REF_AstA_continent <- REF_AsyA_table %>% slice(1:8)
#This code sorts it by country
REF_AsyA_country <- REF_AsyA_table %>% slice(9:115)
  
#Defensive Asylums granted by Nationality from 2010 to 2019
#This code sorts it by years
ref_asyl_grantedD_ByNationality <- read_excel("fy2019_table19d.xlsx")
ref5x <- ref_asyl_grantedD_ByNationality
ref5x <- ref5x[-c(1,2,4,13,104:109), ]
names_ref5x <- ref5x$`Table 19.`
ref5x <- ref5x[,-c(1)]
ref5x <- t(ref5x)
colnames(ref5x) <- names_ref5x
colnames(ref5x)[1] <- "Year"
colnames(ref5x)[98] <- "All Other Countries"
colnames(ref5x)[99] <- "Unknown"
ref5x[ref5x=="X"]<-"NA"
ref5x[ref5x=="D"]<-"witheld"
ref5x[ref5x=="-"]<-0
#This code sorts it by regions
ref_asyl_grantedD_ByNationality <- read_excel("fy2019_table19d.xlsx")
ref5t <- ref_asyl_grantedD_ByNationality
ref5t <- ref5t[-c(1,2,4,13,104:109), ]
names_ref5t <- ref5t[1,]
colnames(ref5t) <- names_ref5t
ref5t <- ref5t[-c(1),]
ref4t[ref5t=="X"]<-"NA"
ref5t[ref5t=="D"]<-"witheld"
ref5t[ref5t=="-"]<-"0"
ref5t[1,1] <- "Total Continent"
ref5t[9,1] <- "Total Country"
REF_AsyD_table <- ref5t
#This code sorts it by continent
REF_AsyD_continent <- REF_AsyD_table %>% slice(1:8)
#This code sorts it by country
REF_AsyD_country <- REF_AsyD_table %>% slice(9:99)


#Naturalization Data
#Naturalizations by Region of Birth from 2010 to 2019
#This code sorts it by years
nat_ByRegionBirth <- read_excel("fy2019_table21d.xlsx")
nat2x <- nat_ByRegionBirth
nat2x <- nat2x[-c(1,2,4,13,216:222), ]
names_nat2x <- nat2x$`Table 21.`
nat2x <- nat2x[,-c(1)]
nat2x <- t(nat2x)
colnames(nat2x) <- names_nat2x
colnames(nat2x)[1] <- "Year"
colnames(nat2x)[98] <- "All Other Countries"
nat2x[nat2x=="X"]<-"NA"
nat2x[nat2x=="D"]<-"witheld"
nat2x[nat2x=="-"]<-0
#This code sorts it by regions
nat_ByRegionBirth <- read_excel("fy2019_table21d.xlsx")
nat2t <- nat_ByRegionBirth
nat2t <- nat2t[-c(1,2,4,13,216:222), ]
names_nat2 <- nat2[1,]
colnames(nat2t) <- names_nat2t
nat2t <- nat2t[-c(1),]
nat2t[nat2t=="X"]<-"NA"
nat2t[nat2t=="D"]<-"witheld"
nat2t[nat2t=="-"]<-"0"
nat2t[1,1] <- "Total Continent"
nat2t[9,1] <- "Total Country"
NAT_Birth_table <- nat2t
#This code sorts it by continent
NAT_Birth_continent <- NAT_Birth_table %>% slice(1:8)
#This code sorts it by country
NAT_Birth_country <- NAT_Birth_table %>% slice(9:211)

#Non-immigrants (Visa Recipients) Data
#Visas granted by Country of Citizenship from 2010 to 2019
#This code sorts it by years
nonimg_ByRegionCitizenship <- read_excel("fy2019_table26.xlsx")
nim1x <- nonimg_ByRegionCitizenship
nim1x <- nim1x[-c(1,2,4,13,210:225), ]
names_nim1x <- nim1x$`Table 26.`
nim1x <- nim1x[,-c(1)]
nim1x <- t(nim1x)
colnames(nim1x) <- names_nim1x
colnames(nim1x)[1] <- "Year"
nim1x[nim1x=="X"]<-"NA"
nim1x[nim1x=="D"]<-"witheld"
nim1x[nim1x=="-"]<-0
#This code sorts it by regions
nonimg_ByRegionCitizenship <- read_excel("fy2019_table26.xlsx")
nim1t <- nonimg_ByRegionCitizenship
nim1t <- nim1t[-c(1,2,4,13,210:225), ]
names_nim1t <- nim1t[1,]
colnames(nim1t) <- names_nim1t
nim1t <- nim1t[-c(1),]
nim1t[nim1t=="X"]<-"NA"
nim1t[nim1t=="D"]<-"witheld"
nim1t[nim1t=="-"]<-"0"
nim1t[1,1] <- "Total Continent"
nim1t[9,1] <- "Total Country"
NIM_Cit_table <- nim1t
#This code sorts it by continent
NIM_Cit_continent <- NIM_Cit_table %>% slice(1:8)
#This code sorts it by country
NIM_Cit_country <- NIM_Cit_table %>% slice(9:205)

#Visas granted by last country of Residence from 2010 to 2019
#This code sorts it by years
nonimg_ByRegionResidence <- read_excel("fy2019_table27d.xlsx")
nim2x <- nonimg_ByRegionResidence
nim2x <- nim2x[-c(1,2,4,13,255:264), ]
names_nim2x <- nim2x$`Table 27.`
nim2x <- nim2x[,-c(1)]
nim2x <- t(nim2x)
colnames(nim2x) <- names_nim2x
colnames(nim2x)[1] <- "Year"
nim2x[nim2x=="X"]<-"NA"
nim2x[nim2x=="D"]<-"witheld"
nim2x[1,1]=2010
#This code sorts it by regions
nonimg_ByRegionResidence <- read_excel("fy2019_table27d.xlsx")
nim2t <- nonimg_ByRegionResidence
nim2t <- nim2t[-c(1,2,4,13,255:264), ]
names_nim1t <- nim2t[1,]
colnames(nim2t) <- names_nim2t
nim2t <- nim2t[-c(1),]
nim2t[nim2t=="X"]<-"NA"
nim2t[nim2t=="D"]<-"witheld"
nim2t[nim2t=="-"]<-"0"
nim2t[1,1] <- "Total Continent"
nim2t[9,1] <- "Total Country"
NIM_Res_table <- nim2t
#This code sorts it by continent
NIM_Res_continent <- NIM_Res_table %>% slice(1:8)
#This code sorts it by country
NIM_Res_country <- NIM_Res_table %>% slice(9:250)

#Aliens and Law Enforcement Data
#'Alien' apprehendings by Nationality from 2010 to 2019
#This code sorts it by years
aliens_appreh_ByRegionNat  <- read_excel("fy2019_table34d.xlsx")
enf2x <- aliens_appreh_ByRegionNat
enf2x <- enf2x[-c(1,2,4,13,180:187), ]
names_enf2x <- enf2x$`Table 34.`
enf2x <- enf2x[,-c(1)]
enf2x <- t(enf2x)
colnames(enf2x) <- names_enf2x
colnames(enf2x)[1] <- "Year"
enf2x[1,1]=2010
enf2x[8,1]=2016
colnames(enf2x)[174] <- "All Other Countries"
enf2x[enf2x=="X"]<-"NA"
enf2x[enf2x=="D"]<-"witheld"
enf2x[enf2x=="-"]<-0
#This code sorts it by regions
aliens_appreh_ByRegionNat  <- read_excel("fy2019_table34d.xlsx")
enf2t <- aliens_appreh_ByRegionNat
enf2t <- enf2t[-c(1,2,4,13,180:187), ]
names_enf2t <- enf2t[1,]
colnames(enf2t) <- names_nenf2t
enf2t <- enf2t[-c(1),]
enf2t[enf2t=="X"]<-"NA"
enf2t[enf2t=="D"]<-"witheld"
enf2t[enf2t=="-"]<-"0"
enf2t[1,1] <- "Total Continent"
enf2t[9,1] <- "Total Country"
ENF_NatA_table <- enf2t
#This code sorts it by continent
ENF_NatA_continent <- ENF_NatA_table %>% slice(1:8)
#This code sorts it by country
ENF_NatA_country <- ENF_NatA_table %>% slice(9:175)

#"Aliens' deemed inadmissible by Nationality from 2010 to 2019
#This code sorts it by years
aliens_inadmin_ByRegionNat <- read_excel("fy2019_table37d.xlsx")
enf5x <- aliens_inadmin_ByRegionNat
enf5x <- enf5x[-c(1,2,4,13,197:201), ]
names_enf5x <- enf5x$`Table 37.`
enf5x <- enf5x[,-c(1)]
enf5x <- t(enf5x)
colnames(enf5x) <- names_enf5x
colnames(enf5x)[1] <- "Year"
colnames(enf5x)[191] <- "All Other Countries"
enf5x[enf5x=="D"]<-"witheld"
enf5x[enf5x=="-"]<-0
#This code sorts it by regions
aliens_inadmin_ByRegionNat <- read_excel("fy2019_table37d.xlsx")
enf5t <- aliens_inadmin_ByRegionNat
enf5t <- enf5t[-c(1,2,4,13,197:201), ]
names_enf5t <- enf5t[1,]
colnames(enf5t) <- names_nenf5t
enf5t <- enf5t[-c(1),]
enf5t[enf5t=="X"]<-"NA"
enf5t[enf5t=="D"]<-"witheld"
enf5t[enf5t=="-"]<-"0"
enf5t[1,1] <- "Total Continent"
enf5t[9,1] <- "Total Country"
ENF_NatI_table <- enf5t
#This code sorts it by continent
ENF_NatI_continent <- ENF_NatI_table %>% slice(1:8)
#This code sorts it by country
ENF_NatI_country <- ENF_NatI_table %>% slice(9:175)

#'Alien' removals by Nationality from 2010 to 2019
#This code sorts it by years
aliens_remov_ByRegionNat <- read_excel("fy2019_table40d.xlsx")
enf8x <- aliens_remov_ByRegionNat
enf8x <- enf8x[-c(1,2,4,13,188:193), ]
names_enf8x <- enf8x$`Table 40.`
enf8x <- enf8x[,-c(1)]
enf8x <- t(enf8x)
colnames(enf8x) <- names_enf8x
colnames(enf8x)[1] <- "Year"
colnames(enf8x)[182] <- "All Other Countries"
enf8x[enf8x=="D"]<-"witheld"
enf8x[enf8x=="-"]<-0
enf8x[7,1]=2016
#This code sorts it by regions
aliens_remov_ByRegionNat <- read_excel("fy2019_table40d.xlsx")
enf8t <- aliens_remov_ByRegionNat
enf8t <- enf8t[-c(1,2,4,13,188:193), ]
names_enf5t <- enf8t[1,]
colnames(enf8t) <- names_nenf8t
enf8t <- enf8t[-c(1),]
enf8t[enf8t=="X"]<-"NA"
enf8t[enf8t=="D"]<-"witheld"
enf8t[enf8t=="-"]<-"0"
enf8t[1,1] <- "Total Continent"
enf8t[9,1] <- "Total Country"
ENF_NatR_table <- enf8t
#This code sorts it by continent
ENF_NatR_continent <- ENF_NatR_table %>% slice(1:8)
#This code sorts it by country
ENF_Natr_country <- ENF_NatR_table %>% slice(9:175)


#'Alien' removals by Nationality and Crime Status from 2010 to 2019
aliens_remov_byCrimS_RegionNat <- read_excel("fy2019_table41d.xlsx")
enf9 <- aliens_remov_byCrimS_RegionNat

#'Alien' removals by Nationality and Crime Status in 2010
enf9_2010 <- enf9[1:4]
enf9_2010 <- enf9_2010[-c(1:5,14,176:181), ]
colnames(enf9_2010) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2010[enf9_2010=="D"]<-"witheld"
enf9_2010[enf9_2010=="-"]<-"0"
enf9_2010[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2011
enf9_2011 <- enf9[1:7]
enf9_2011 <- enf9_2011[-c(1:5,14,176:181),-c(2:4)]
colnames(enf9_2011) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2011[enf9_2011=="D"]<-"witheld"
enf9_2011[enf9_2011=="-"]<-"0"
enf9_2011[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2012
enf9_2012 <- enf9[1:10]
enf9_2012 <- enf9_2012[-c(1:5,14,176:181),-c(2:7)]
colnames(enf9_2012) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2012[enf9_2012=="D"]<-"witheld"
enf9_2012[enf9_2012=="-"]<-"0"
enf9_2012[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2013
enf9_2013 <- enf9[1:13]
enf9_2013 <- enf9_2013[-c(1:5,14,176:181),-c(2:10)]
colnames(enf9_2013) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2013[enf9_2013=="D"]<-"witheld"
enf9_2013[enf9_2013=="-"]<-"0"
enf9_2013[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2014
enf9_2014 <- enf9[1:16]
enf9_2014 <- enf9_2014[-c(1:5,14,176:181),-c(2:13)]
colnames(enf9_2014) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2014[enf9_2014=="D"]<-"witheld"
enf9_2014[enf9_2014=="-"]<-"0"
enf9_2014[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2015
enf9_2015 <- enf9[1:19]
enf9_2015 <- enf9_2015[-c(1:5,14,176:181),-c(2:16)]
colnames(enf9_2015) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2015[enf9_2015=="D"]<-"witheld"
enf9_2015[enf9_2015=="-"]<-"0"
enf9_2015[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2016
enf9_2016 <- enf9[1:22]
enf9_2016 <- enf9_2016[-c(1:5,14,176:181),-c(2:19)]
colnames(enf9_2016) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2016[enf9_2016=="D"]<-"witheld"
enf9_2016[enf9_2016=="-"]<-"0"
enf9_2016[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2017
enf9_2017 <- enf9[1:25]
enf9_2017 <- enf9_2017[-c(1:5,14,176:181),-c(2:21)]
colnames(enf9_2017) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2017[enf9_2017=="D"]<-"witheld"
enf9_2017[enf9_2017=="-"]<-"0"
enf9_2017[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2018
enf9_2018 <- enf9[1:28]
enf9_2018 <- enf9_2018[-c(1:5,14,176:181),-c(2:24)]
colnames(enf9_2018) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2018[enf9_2018=="D"]<-"witheld"
enf9_2018[enf9_2018=="-"]<-"0"
enf9_2018[168,1] <- "All Other Countries"

#'Alien' removals by Nationality and Crime Status in 2019
enf9_2019 <- enf9[1:31]
enf9_2019 <- enf9_2019[-c(1:5,14,176:181),-c(2:27)]
colnames(enf9_2011) <- c("Country","Total Alien Removals","Criminal Alien Removed","Non-Criminal Alien Removed")
enf9_2019[enf9_2019=="D"]<-"witheld"
enf9_2019[enf9_2019=="-"]<-"0"
enf9_2019[168,1] <- "All Other Countries"

```
## 1. Table Creation
 I then made simple tables with the different datasets that included State of Residence. I also used the last data set in the "Region" group, enf9 which had the number of alien removals by country and by criminal status for the fiscals years 2010 to 2019. After splitting the data by year, I made a graph for each year.

```{r Table Creation, eval=FALSE}

LPR_state_table %>%gt() %>%
  tab_header( title = "PERSONS OBTAINING LAWFUL PERMANENT RESIDENT STATUS BY STATE OR TERRITORY OF RESIDENCE: FISCAL YEARS 2010 TO 2019")

NAT_state_table %>%gt() %>%
  tab_header( title = "NATURALIZATIONS BY STATE OR TERRITORY OF RESIDENCE: FISCAL YEARS 2010 TO 2019")

enf9_2010 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2010")

enf9_2011 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2011")

enf9_2012 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2012")

enf9_2013 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2013")

enf9_2013 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2014")

enf9_2015 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2015")

enf9_2016 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2016")

enf9_2017 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2017")

enf9_2018 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2018")

enf9_2019 %>% gt() %>% 
  tab_header(title="Alien Removals by Country and Criminal Status in 2019")
  
```
## 1. Graph Creation
I then made graphs of the datasets that spanned over 20 years


```{r Graph Creation, eval=FALSE}

#Lawful Permanent Residents Data
ggplot(data=LPR_years, aes(x=Year,y=`People Obtaining LPR Status`)) + 
  geom_point(color="yellow", size=3) +
  xlab("Year") +
  ylab("People Obtaining LPR Status") +
  ggtitle("PERSONS OBTAINING LAWFUL PERMANENT RESIDENT STATUS: FISCAL YEARS 2000 TO 2019") +
  theme_dark() 


#Refugees and Asylees Data
ggplot(data=REF, aes(x=Year,y=`Refugee Arrivals`)) + 
  geom_point(color="yellow", size=3) +
  xlab("Year") +
  ylab("Refugee Arrivals") +
  ggtitle("REFUGEE ARRIVALS: FISCAL YEARS 2000 TO 2019") +
  theme_dark() 

ggplot(data=REF, aes(x=Year,y=`Total Asylums Granted`)) + 
  geom_point(color="yellow", size=3) +
  xlab("Year") +
  ylab("Total Asylums Granted") +
  ggtitle("ASYLUM STATUS ARRIVALS: FISCAL YEARS 2000 TO 2019") +
  theme_dark() 


#Aliens and Law Enforcement Data
ggplot(data=ENF_years, aes(x=Year,y=`Aliens Apprehended`)) + 
  geom_point(color="yellow", size=3) +
  xlab("Year") +
  ylab("Total") +
  ggtitle("ALIENS APPREHENDED: FISCAL YEARS 2000 TO 2019") +
  theme_dark() 

ggplot(data=ENF_years, aes(x=Year,y=`Aliens Determined Inadmissible`)) + 
  geom_point(color="yellow", size=3) +
  xlab("Year") +
  ylab("Total") +
  ggtitle("ALIENS DETERMINED INADMISSIBLE: FISCAL YEARS 2000 TO 2019") +
  theme_dark() 

ggplot(data=ENF_years, aes(x=Year,y=`Aliens Removals`)) + 
  geom_point(color="yellow", size=3) +
  xlab("Year") +
  ylab("Total") +
  ggtitle("ALIEN REMOVALS: FISCAL YEARS 2000 TO 2019") +
  theme_dark() 

ggplot(data=ENF_years, aes(x=Year,y=`Aliens Returned`)) + 
  geom_point(color="yellow", size=3) +
  xlab("Year") +
  ylab("Total") +
  ggtitle("ALIEN RETURNS: FISCAL YEARS 2000 TO 2019") +
  theme_dark() 
```

## 1. Analysis
  While looking at the graphs, we cans see that there was a sharp decline in immigration flux to the United States in the period right after 2001 and the presidential election of 2004. However there was also a decline of immigration flux after the 2016 presidential election, from which we cna infer that republican administrations could be more strict on immigration.
  Something else that we can observe from both the graphs and the tables is that when it comes to 'Aliens" apprehending, removals, and returns, there is no clear trend as the numbers fluctuate throughout the years and mid presidencies. It seems that this trend is more related to each president and his policy, rather than a political party ideology
  Finally, when it comes to accepting refugees and granting asylum status, it seems like numbers went way down after 2001, then gradually rose throughout the years and then significantly dropped around 2017-2018, whihc owuld have been when Donal Trump took office and started enacting his executive orders

## 1. Conclusion
  It seems that when it comes to immigration towards regarding the United States, the influx does not really decrease based on the presidency, it actually does seem to increase for some country of provenance as the years go by. What varies based on the presidency, is the number of Aslums granted and most importantly the treatment of 'Aliens' (or immigrants deemed 'non-legal'). 
  
  As we could have imagined, a traumatic event for the United States like 9/11 will affect those numbers in a negative way as xenophobia might run rampant and in the name of security, a higher scrutiny might be applied to all. But what is very surprising is that attitudes towards 'aliens' is not a partisan issue: as in both Republican presidents and the only Democrat President in the data have had similar numbers when it comes to the treatment of those "non-legal" immigrants. 
  
  However, it might not be fair to take the Data for one Democrat president and consider as a general trend for the entire party. Since we cannot utlize the data from previous years as it is missing and would not exactly correlate as it is not in the same context, the best solution would be take in account the data for the next presidency (which would be Democrat) and see whether this Democrat president follows in the footsteps of his direct predecessor, his last Democrat predecessor or instead follows a whole other trend. Those numbers will obviously be skewed because of the pandemic but we could try to adapt the data with a  model to see what it would have been without the pandemic.